Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming

### 2022-02-14 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
update step-by-step user guide

\subsubsection{Numerical equations}

@@

\subsection{Diagnosis of the approximation} \label{subsec_diagnosis}

@@

\subsubsection{User guide with the steps to perform diagnosis}

\paragraph*{Assumption of the diagnosis tool}: Under inaccurate HSGP approximation, the estimated lengthscale $\hat{l}$ is smaller than the minimum length-scale $l^*$ determined by $m$ and $c$ in Figure~\ref{fig6_7_relationships}. Thus the diagnostic pointing towards an accurate approximation is:
%
\begin{align}\label{eq_diagnostic}
&\hat{l} \pm 0.02 \geq l^*
\end{align}

\paragraph*{User-guide with the steps to perform diagnosis}:

\begin{enumerate}
\item[Step 1.] Make a first guess on the lengthscale $l_1$ of the function to be learned. We recommend a large lengthscale, around 1.
\item[] \hspace{-1cm} \textit{(Initial iteration, $i=1$)}
\item[Step 2.] Obtain the valid minimum boundary factor $c_1$ determined by $l_1$ by using (\ref{eq_c_vs_l_QE}).
\item[Step 3.] Obtain the valid mimimum number of basis functions $m_1$ determined by $l_1$ and $c_1$ by using (\ref{eq_m_l_QE}). Notice that $l_1$ can also be read as the minimum lengthscale that can be accurately fitted determined by $m_1$ and $c_1$ by using (\ref{eq_m_l_QE}).
\item[Step 4.] Fit the HSGP model and assess residuals: compute \textit{rmse} (Root mean square error), $R^2$ (Coefficient of variation) and \textit{elpd} (Expected log predicitve density).
\item[Step 5.] Check the diagnostic of whether $\, \hat{l_1} \pm 0.02 \geq l_1$ (\ref{eq_diagnostic}).
	\begin{enumerate}
    \item If the diagnostic is TRUE, the HSGP model approximation must be close to be sufficiently accurate. Then user can continue with step 6.
       	\begin{enumerate}
    	\item[] \hspace{-1cm} \textit{(Next iterations, $i>1$)}
		\item[Step 6.] Set $\, m_i = m_{i-1} + 10$.
		\item[Step 7.] Obtain the valid minimum boundary factor $c_i$ by using (\ref{eq_c_vs_l_QE}), with $c(\hat{l}_{i-1})$.
		\item[Step 8.] Obtain the mimimum lengthscale $l_i$ that can be accurately fitted determined by $m_i$ and $c_i$ by using (\ref{eq_m_l_QE}).
		\item[Step 9.] Fit the HSGP model and
			\begin{enumerate}
        	\item[Step 9.1.] Check whether $\, \hat{l}_i \pm 0.02 \geq l_i$ (\ref{eq_diagnostic}),
        	\item[Step 9.2.] Check stability in $\hat{l}_i$, \textit{rmse}, $R^2$ and \textit{elpd} relative to previous iteration.
			\end{enumerate}
		\item[Step 10.] If the verifications in steps 9.1 and 9.2 are TRUE, the HSGP model approximation should be sufficiently accurate, and diagnosis ends here. Otherwise, repeat steps 6-10 and update parameters.
    	\end{enumerate}
     \item If the diagnostic is FALSE, the HSGP model approximation can not be sufficiently accurate. Then user has to move to step 11.
     	\begin{enumerate}
		\item[Step 11.] Set $\, l_i = \hat{l}_{i-1}$.
		\item[Step 12.] Repeat steps 2-5 and update parameters.
		\end{enumerate}
    \end{enumerate}
\end{enumerate}

\subsection{Performance analysis of the diagnosis}

>>>>

\subsubsection{Numerical equations}
\label{num_equations}

@@

\subsection{Diagnostics of the approximation} \label{subsec_diagnostics}

@@

\subsubsection{A step-by-step user-guide to apply the diagnostics}

Based on the above proposed diagnostics, we obtain a simple step-by-step procedure that users can follow in order to obtain an accurate HSGP approximation. The procedure is split into two phases, Phase A and B, which have to be completed consequtively.
%In the following, we will refer to the Equations from Section XXX for the double-exponential kernel, but the same procedure works for other kernels by using the corresponding Equations provided in the same Section.

\paragraph*{Phase A:}

\begin{enumerate}
\item[Step A1.] Make an initial guess on true the lengthscale $l_1$ of the function to be learned. We recommend a large lengthscale, about $l_1 \approx 1$. Set the iteration index to $i=1$.
%\item[] \hspace{-1cm} \textit{(Initial iteration, $i=1$)}
\item[Step A2.] Obtain the minimum valid boundary factor $c_i$ determined by $l = l_i$ for the given kernel as per Section \ref{num_equations} 
\item[Step A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $l = l_i$ and $c = c_i$ for the given kernel as per Section \ref{num_equations}. 
\item[Step A4.] Fit an HSGP model using $m_i$ and $c_i$ and ensure convergence of the MCMC chains.
\item[Step A5.] Perform the lengthscale diagnostic by checking if $\hat{l}_i \pm 0.02 \geq l_i$.

If the diagnostic is FALSE, the HSGP approximation is not yet sufficiently accurate. Set $l_{i+1} = \hat{l}_{i}$, increase the iteration index $i = i + 1$, and go back to Step A2.

If the diagnostic is TRUE, the HSGP approximation is close to be sufficiently accurate. Continue with Phase B.

\end{enumerate}

\paragraph*{Phase B:}

\begin{enumerate}

\item[Step B1.] For the current HSGP model, compute measures of predictive accuracy, for example, root mean square error (RMSE), coefficient of determination ($R^2$), or expected log predicitve density (ELPD).

\item[Step B2.] Set $m_{i+1} = m_{i} + 10$ and increase the iteration index $i = i + 1$.

\item[Step B3.] Obtain the minimum valid boundary factor $c_i$ determined by $l = \hat{l}_{i-1}$ for the given kernel as per Section \ref{num_equations}.

\item[Step B4.] Obtain the mimimum valid lengthscale $l_i$ that can be accurately estimated by $m = m_i$ and $c = c_i$ for the given kernel as per Section \ref{num_equations}.

\item[Step B5.] Fit an HSGP model using $m_i$ and $c_i$ and ensure convergence of the MCMC chains.

\item[Step B6.] Perform the lengthscale diagnostic by checking if $\hat{l}_i \pm 0.02 \geq l_i$.

Check the stability of both $\hat{l}_i$ and the measures of predictive accuracy relative to the previous iteration.

If all the stability checks succeeed, the HSGP approximation of the latest model should be sufficiently accurate and the procedure ends here.

Otherwise, go back to Step B1.
\end{enumerate}

\subsection{Performance analysis of the diagnostics}
